---
title: "Simulate Integrated CIR process"
author: "Michael B Gordy"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r generateBibliography, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
require("knitcitations")
cleanbib()
options("citation_format" = "pandoc")
# read.bibtex(file = "bibliography.bib")
```


Let $Y[t]$ denote the integral of a CIR process from time 0 to $t$.  This integral arises frequently in applications to finance.  The function `rCIRintegrated()` samples $Y[t]$ conditional on $X[0]$ and $X[t]$ using the (almost) exact sampling method of @glassermanKim2011. In this vignette, we demonstrate that this routine works well relative to a traditional quadrature method.

Let $H$ denote the time horizon, and $K$ the number of grid points for quadrature. We define a function `rCIRtransition2.trapezoid()` that returns a list $(X[H], Y[H])$ by sampling $X[t]$ at $t=(H/K),2(H/K),\ldots,H$. At each grid point, $Y[t]$ is calculated by the trapezoid approximation.

```{r}
library(basicAffineProcess)
rCIRtransition2.trapezoid <- function(X0,mu,kappa,sigma,horizon,K) {
  xpath <- rCIRpath(X0,mu,kappa,sigma,horizon,ticks=K)$Xt
  XH <- xpath[K+1]
  YH <- sum(xpath)-(X0+XH)/2
  return(list(Xt=XH,Yt=YH))
}

```

We fix our parameter values. 

```{r}
mu <- 1.75
kappa <- 0.35
sigma <- 0.25
X0 <- 0.64      # for simplicity we fix a starting value for X[0]

horizon <- 1
K <- 40  
simtrials <- 500

```

Now let's simulate.  For each trial *i*, we use `rCIRtransition2.trapezoid()` to draw $(X^i[H],Y_{trp}^i[H])$, and then use `rCIRintegrated()` in the package to sample $Y_{gh}^i[H]$ conditional on the $X^i[H]$ just drawn.  Here we use subscripts *trp* to denote the trapezoid algorithm and *gk* to denote Glasserman-Kim.

```{r}
XH <- YH_gk <- YH_trp <- rep(0,simtrials)
for (i in 1:simtrials) {
  pathi <- rCIRtransition2.trapezoid(X0,mu,kappa,sigma,horizon,K)
  YH_trp[i] <- pathi$Yt
  XH[i] <- pathi$Xt
  YH_gk[i] <- rCIRintegrated(X0,XH[i],mu,kappa,sigma,horizon)
}
```

If the Glasserman-Kim routine works correctly, then a QQ plot will show that the two samples of $Y[H]$ have the same unconditional distribution.  The dotted line represents the $y=x$ line.

```{r fig.width=7, fig.height=7}
qqplot(YH_trp,YH_gk, main="Q-Q Plot", xlab='Y[H] by trapezoid rule',ylab='Y[H] by G-K')
yrange <- c(min(YH_trp),max(YH_trp))
lines(yrange,yrange,col=4,lty=3)
```

Not bad!  Recall that the trapezoid rule yields a sampler that is far from an exact, any small discrepancy in the distributions probably should be intepreted in favor of the Glasserman-Kim method.

## References


